### 2018-08-07 main.tex

\documentclass[format=sigconf, screen=true]{acmart}

%-------------- Customization of acmsmall package begins ----------------
\settopmatter{printacmref=false}%removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} 	% removes footnote with conference information in first column
\fancyfoot[LE,RO]{}									%removes footer with publication date
\fancypagestyle{firstpagestyle}{\fancyhf{}}			%removes footer with publication date on first page
\fancyhead[LE,RO]{\thepage}		%removes article number from header and leaves only page number
\makeatletter					%removes the black rectangle with article number
\renewcommand\@folioblob{}
\makeatother
%-------------- Customization of acmsmall package ends ----------------

\usepackage{thmtools}
\declaretheorem[style=acmdefinition,qed=$\vartriangleleft$,sibling=theorem,name=Example]{myexample}
\declaretheorem[style=acmplain,sibling=theorem]{remark}

%Import personal commands 
\usepackage{commandsrm}
%Load cleveref last
\usepackage[capitalize]{cleveref}

\usepackage{epigraph}
\setlength\epigraphwidth{8cm}
\setlength\epigraphrule{0pt}

% DOI
%\acmDOI{0000001.0000001}

\begin{document}
\title{Clever Title}

\author{Alessandro Arlotto} 
\affiliation{Duke University}
\author{Siddhartha Banerjee}
\affiliation{Cornell University}
\author{Itai Gurvich}
\affiliation{Cornell University}
\author{Alberto Vera}
\affiliation{Cornell University}

\keywords{Online Stochastic Optimization, Prophet Inequalities, Approximate Dynamic Programming, Network Revenue Management, Online Packing.}
\begin{abstract}
\input{sec_abstract.tex}
\end{abstract}

\maketitle

\section{Introduction}

\input{sec_intro.tex}

\section{Relaxed Bellman Equations}
\input{sec_bellman.tex}

\section{Probing}
\input{sec_probing.tex}

\section{Stochastic Knapsack With Frequentist Learning}
\input{sec_knapsack.tex}

\section{Conclusions}
\input{sec_conclu.tex}

%% Bibliography
\bibliographystyle{plain}
\bibliography{biblio}

\end{document}