### 2018-08-07 sec_knapsack.tex

\paragraph{Problem Statement}
Items with known reward $r$ and random weight $W$ can be placed in a knapsack of capacity $B$.
At every time $t$, an arrival $j\in[n]$ is drawn with some known distribution, which generates a known reward of $r_j$.
The weight $W_j$ is a random variable, which is revealed only after the decision of accept/reject has been made.
The support of $W_j$ is known, but not the distribution.
An arrival can be accepted only if the weight is a.s.\ smaller than the remaining capacity.
At the end of each period, we observe the realization of both accepted and rejected items.

\off has access to the distribution of weights, \emph{but not to the realizations}.
Let $w_j\defeq \E[W_j]$ and $\Wt_j$ be the empirical average of the weights observed.
We assume that, before the process starts, we are given one sample of each weight type to initialize $\hat W_{jt}$.

Let $N_j(t)$ be the number of samples type $j$ observed at the beginning of $t$.
By definition we have $N_j(t)=Z_j(\T)-Z_j(t)+1$.
Following the same arguments, we can show that the relaxation in \cref{eq:knapsack_lp2} satisfies the initial condition.
\begin{equation}\label{eq:knapsack_lp2}
\begin{array}{rrll}
\varphi(t,b,\xit) \; = \; \max & \multicolumn{3}{l}{\sum_{j}r_{j}x_j} \\
\text{s.t.}& \sum_{j} w_jx_j  &\leq b  \\
&  x_j & \leq Z_j(t)  & j\in [n] \\
& x&\geq 0.
\end{array}
\end{equation}

Define as $\hat \varphi(t,b)$ the LP obtained from \cref{eq:knapsack_lp2} by replacing $Z_j(t)$ and $w_j$ by $\mu_j(t)\defeq\E[Z_j(t)]$ and $\Wt_j$, respectively.
Let $\sigma_j\in[n]$ be the rank of $j$ w.r.t.\ the ratio $r_j/w_j$, breaking ties lexicographically. 
For example, if $\sigma_j=1$, it means that $j$ is the most desirable item.
Similarly, $\sigmat_j$ denotes the rank of $j$ w.r.t.\ the ratio $r_j/\Wt_j$.

\begin{lemma}
Assume there are two items ($n=2$)  and $\abs{w_1-w_2}=\Delta>0$.
Let $\Xts,\Xt$ be the solutions to $\varphi,\hat\varphi$ respectively, then
\[
\Pr[\Xt_j\geq \mu_j(t)/2, \Xts<1] \leq STUFF, \quad 
\Pr[\Xt_j< \mu_j(t)/2, \Xts>Z_j(t)-1]\leq STUFF
\]
\end{lemma}
\begin{proof}
Let us bound the first probability by partitioning the space as
\[
\Pr[\Xt_j\geq \mu_j(t)/2, \Xts<1] \leq \Pr[\Xt_j\geq \mu_j(t)/2, \Xts<1, \sigma_j=\sigmat_j] + \Pr[\sigma_j\neq\sigmat_j].
\]
The term $\Pr[\sigma_j\neq\sigmat_j]$ decays with the number of samples $N_1(t),N_2(t)$.
We now bound the other term.

Say $j=1$ and let us study the case $\sigma_1=1$ first.
We can write $\Xts_1 = Z_1(t)\land \frac{b}{w_1}$ and $\Xt_1 = \mu_1(t)\land \frac{b}{\Wt_1}$.
Since we are assuming $\Xt_1\geq \mu_1(t)/2$, it follows that $\frac{b}{\Wt_1}\geq \mu_1(t)/2$.
On the other hand, $\Xts_1<1$ necessitates $\frac{b}{w_1}<1$ or $Z_1(t)=0$.
We can use these two inequalities to conclude
\[
\Pr[\Xt_j\geq \mu_j(t)/2, \Xts<1, \sigma_j=\sigmat_j=1] \leq \Pr[ w_1> \Wt_1 \mu_1(t)/2 \text{ or }  Z_1(t)=0].
\]
The probability of both events decay with $\mu_1(t)$.

Finally, let us study the case $\sigma_1=2$.
We can write $\Xts_2 = Z_2(t)\land \frac{b}{w_2}$, $\Xt_2 = \mu_2(t)\land \frac{b}{\Wt_2}$, $\Xts_1 = Z_1(t)\land \frac{b-\Xts_2w_2}{w_1}$ and $\Xt_1 = \mu_1(t)\land \frac{b-\Xt_2\Wt_2}{\Wt_1}$.
The conditions $\Xt_1\geq \mu_1(t)/2$ and $\Xts_1<1$ imply
\begin{align*}
& \Wt_1\frac{\mu_1(t)}{2}+\Xt_2\Wt_2\leq b < w_1+\Xts_2w_2 \quad \text{ or } \quad Z_1(t)=0\\
&\Longrightarrow \Wt_1\frac{\mu_1(t)}{2}< w_1+ (Z_2(t)w_2)\land b - (\mu_2(t)\Wt_2)\land b \quad \text{ or } \quad Z_1(t)=0 \\
&\Longrightarrow \Wt_1\frac{\mu_1(t)}{2}< w_1+ \abs{Z_2(t)w_2 -\mu_2(t)\Wt_2} \quad \text{ or } \quad Z_1(t)=0.
\end{align*}
For the last implication we used the fact that $x\land b - y\land b \leq \abs{x-y}$ holds for any reals $x,y,b$.
We conclude by observing that both events have probability decaying with $\mu_1(t),N_1(t),N_2(t)$.
\end{proof}